---
title: "品質管制 Homework 11"
header-includes:
  - \usepackage{xeCJK}
  - \setCJKmainfont{標楷體}
output: 
  pdf_document:
    latex_engine: xelatex
    number_sections: true
author: "110024516 統研碩一 邱繼賢"
---

**5.20**  
```{r include=FALSE}
eta_1 = function(en, lambda, u) {
    if (en < -u) {
        return(en+(1-lambda)*u)
    } else if (en > u) {
        return(en-(1-lambda)*u)
    } else {
        return(lambda*en)
    }
}

eta_2 = function(en, lambda, u) {
    if ((abs(en) < u) | (abs(en) == u)) {
        return(en*(1-(1-lambda)*(1-(en/u)^2)^2))
    } else {
        return(en)
    }
}

Xi = c(0.0, -2.3, 0.6, -1.0, -0.2, 0.3, 0.5, -0.4, 0.3, -0.5,
       -0.9, -0.5, 1.0, -0.9, 1.5, 1.4, 0.0, 2.2, 2.7, 0.7, -0.4, -0.8, -0.1, 0.0, 0.3, -1.0, 2.5, 1.3, 0.3, -0.3)
```

**(ii)**  
$ARL_0=100\ ,\ \eta(e_n)=\eta_1(e_n)\ ,\ (\delta_1,\delta_2)=(1.0,4)\ ,\ \text{and}\ v=0.05$  
$\text{by Table 5.6}\ \Rightarrow\ h=0.7874\ ,\ \lambda=0.1813\ ,\ u=2.5752$  
The charting statistic of the AEWMA chart is defined by
$$
\begin{aligned}
&A_n\ =\ A_{n-1}\ +\ \eta_1(e_n)\ ,\ \ \text{where }\ A_0\ =\ \mu_0\ =\ 0\ ,\ e_n\ =\ X_n\ -\ A_{n-1}
\end{aligned}
$$
and 
$$
\eta_1(e_n)\ =\ 
\left\{
\begin{aligned}
e_n\ +\ (1-\lambda)\ u\ \ \ ,\ \ &\text{if}\ \ e_n<-u\\
\lambda\ e_n\ \ \ ,\ \ &\text{if}\ \ |e_n|\leq u\\
e_n\ -\ (1-\lambda)\ u\ \ \ ,\ \ &\text{if}\ \ e_n>u
\end{aligned}
\right.
$$
The results of charting statistics are shown as below :

```{r echo=FALSE}
library(knitr)
lambda2 = 0.1813
u2 = 2.5752
h2 = 0.7874

An_2 = c()
en_2 = c()
en_2[1] = Xi[1]-0
An_2[1] = 0+eta_1(en_2[1],lambda2,u2)

for (i in 2:30) {
    en_2[i] = Xi[i] - An_2[i-1]
    An_2[i] = An_2[i-1] + eta_1(en_2[i],lambda2,u2)
}

U2 = 0+h2
L2 = 0-h2

charting_stat = data.frame(X = Xi, 
                           en = en_2, 
                           An = An_2)

kable(charting_stat, col.names = c("$X_n$", "$e_n$", "$A_n$"), digits = 4)
```

The chart signals if $|A_n\ -\ \mu_0|\ >\ h$
$$
\Rightarrow
\left\{
\begin{aligned}
&U\ =\ \mu_0\ +\ h\ =\ 0\ +\ 0.7874\ =\ 0.7874\\
&L\ =\ \mu_0\ -\ h\ =\ 0\ -\ 0.7874\ =\ -0.7874
\end{aligned}
\right.
$$

```{r echo=FALSE}
par(mfrow = c(1,1), mar=c(5,4,4,5))
plot(1:30, An_2, type="o",lty=1,pch=16,xlab="n",ylab=expression(A[n]),
     mgp=c(2,1,0),cex = 0.8, ylim = c(-1, 1))
lines(1:30, rep(U2,30), lty = 2, cex = 0.8)
lines(1:30, rep(L2,30), lty = 2, cex = 0.8)
axis(4, at = c(U2,L2), labels = c("U=0.7874","L=-0.7874"), las = 2)
```

The AEWMA chart detects mean shift at the 19th time point.


\newpage
**(iv)**  
$ARL_0=100\ ,\ \eta(e_n)=\eta_2(e_n)\ ,\ (\delta_1,\delta_2)=(0.25,4)\ ,\ \text{and}\ v=0.05$  
$\text{by Table 5.6}\ \Rightarrow\ h=0.3542\ ,\ \lambda=0.0188\ ,\ u=12.6145$  
The charting statistic of the AEWMA chart is defined by
$$
\begin{aligned}
&A_n\ =\ A_{n-1}\ +\ \eta_2(e_n)\ ,\ \ \text{where }\ A_0\ =\ \mu_0\ =\ 0\ ,\ e_n\ =\ X_n\ -\ A_{n-1}
\end{aligned}
$$
and 
$$
\eta_2(e_n)\ =\ 
\left\{
\begin{aligned}
e_n\ [\ 1-(1-\lambda)(1-(\frac{e_n}{u})^2)^2\ ]\ \ \ ,\ \ &\text{if}\ \ |e_n|\leq u\\
e_n\ \ \ ,\ \ &\text{otherwise}
\end{aligned}
\right.
$$
The results of charting statistics are shown as below :

```{r echo=FALSE}
lambda4 = 0.0188
u4 = 12.6145
h4 = 0.3542

An_4 = c()
en_4 = c()
en_4[1] = Xi[1]-0
An_4[1] = 0+eta_2(en_4[1],lambda4,u4)

for (i in 2:30) {
    en_4[i] = Xi[i] - An_4[i-1]
    An_4[i] = An_4[i-1] + eta_2(en_4[i],lambda4,u4)
}

U4 = 0+h4
L4 = 0-h4

charting_stat4 = data.frame(X = Xi, 
                            en = en_4, 
                            an = An_4)
kable(charting_stat4, col.names = c("$X_n$", "$e_n$", "$A_n$"), digits = 4)

```

The chart signals if $|A_n\ -\ \mu_0|\ >\ h$
$$
\Rightarrow
\left\{
\begin{aligned}
&U\ =\ \mu_0\ +\ h\ =\ 0\ +\ 0.3542\ =\ 0.3542\\
&L\ =\ \mu_0\ -\ h\ =\ 0\ -\ 0.3542\ =\ -0.3542
\end{aligned}
\right.
$$

```{r echo=FALSE}
par(mfrow = c(1,1), mar=c(5,4,4,5))
plot(1:30, An_4, type="o",lty=1,pch=16,xlab="n",ylab=expression(A[n]),
     mgp=c(2,1,0),cex = 0.8, ylim = c(-1, 1))
lines(1:30, rep(U4,30), lty = 2, cex = 0.8)
lines(1:30, rep(L4,30), lty = 2, cex = 0.8)
axis(4, at = c(U4,L4), labels = c("U=0.3542","L=-0.3542"), las = 2)
```

The AEWMA chart detects mean shift at the 19th time point.

```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```

