---
title: "品質管制 Homework 7"
header-includes:
  - \usepackage{xeCJK}
  - \setCJKmainfont{標楷體}
output: 
  pdf_document:
    latex_engine: xelatex
    number_sections: true
author: "110024516 統研碩一 邱繼賢"
date: "2021年11月13日"
---
\textbf{4.13}  
```{r include=FALSE}
compute_k = function(sigma_0, sigma_1) {
    return((2*log(sigma_0/sigma_1))/((sigma_0/sigma_1)^2-1))
}
```

\textbf{(i)}$\ \sigma_0\ =\ 1,\ \sigma_1\ =\ 2,\ ARL_0\ =\ 370$  
To detect the upward variance shift, the CUSUM chart is:
\[
\begin{aligned}
&C_n^+=max(0\ ,\ C_{n-1}^+\ +\ (\frac{X_n-\mu_0}{\sigma_0})^2\ -\ k^+)\ ,\ \text{where}\ C_0^+\ =\ 0\\
&\text{and}\ k^+\ =\ \frac{2\ log(\sigma_0/\sigma_1)}{(\sigma_0/\sigma_1)^2-1}\ =\ 1.8484\\
&\text{This chart gives a signal of upward variance shift if}\ C_n^+>h_U\\
&\text{Where}\ h_U\ =\ 8.97\ \text{is computed by}\ ANYGETH.EXE\\
&\therefore\ \text{The optimal values of}\ (k^+,h_U)\ =\ (1.848,8.97)
\end{aligned}
\]
```{r include=FALSE}
k_plus1 = compute_k(1, 2)
h_u1 = 8.970
```


\textbf{(ii)}$\ \sigma_0\ =\ 1,\ \sigma_1\ =\ 0.5,\ ARL_0\ =\ 370$  
To detect the downward variance shift, the CUSUM chart is:
\[
\begin{aligned}
&C_n^-=min(0\ ,\ C_{n-1}^-\ +\ (\frac{X_n-\mu_0}{\sigma_0})^2\ -\ k^-)\ ,\ \text{where}\ C_0^-\ =\ 0\\
&\text{and}\ k^-\ =\ \frac{2\ log(\sigma_0/\sigma_1)}{(\sigma_0/\sigma_1)^2-1}\ =\ 0.4621\\
&\text{This chart gives a signal of downward variance shift if}\ C_n^-<h_L\\
&\text{Where}\ h_L\ =\ -2.843\ \text{is computed by}\ ANYGETH.EXE\\
&\therefore\ \text{The optimal values of}\ (k^-,h_L)\ =\ (0.4621,-2.843)
\end{aligned}
\]

```{r include=FALSE}
k_minus2 = compute_k(1, 0.5)
h_l2 = -2.843
```


\textbf{(iii)}$\ \sigma_0\ =\ 0.5,\ \sigma_1\ =\ 1,\ ARL_0\ =\ 370$  
To detect the upward variance shift, the CUSUM chart is:
\[
\begin{aligned}
&C_n^+=max(0\ ,\ C_{n-1}^+\ +\ (\frac{X_n-\mu_0}{\sigma_0})^2\ -\ k^+)\ ,\ \text{where}\ C_0^+\ =\ 0\\
&\text{and}\ k^+\ =\ \frac{2\ log(\sigma_0/\sigma_1)}{(\sigma_0/\sigma_1)^2-1}\ =\ 1.8484\\
&\text{This chart gives a signal of upward variance shift if } C_n^+>h_U\\
&\text{Where}\ h_U\ =\ 8.906135\ \text{is computed by the bisection simulation}\\
&and\ the\ last\ ARL_0\ =\ 369.67\\
&\therefore\ \text{The optimal values of}\ (k^+,h_U)\ =\ (1.848,8.906135)
\end{aligned}
\]

```{r include=FALSE}
k_plus3 = compute_k(0.5, 1)
h_u3 = 2.243
```



\textbf{4.14}  
計算$k^-\ =\ \frac{2\ log(\sigma_1/\sigma_0)\ \sigma_0^2\ \sigma_1^2}{\sigma_1^2\ -\ \sigma_0^2}\ =\ \frac{2\ log(0.5/1)\ 0.25}{0.25\ -\ 1}\ =\ 0.4621$，再使用$ANYGETH.EXE$軟體套件計算$h_L\ =\ -0.681$
```{r}
k_minus = (1*0.25*log(0.25/1))/(0.25-1)
h_l = -0.681
```

匯入資料並計算\[C_n^-\ =\ min(0, C_{n-1}^-\ +\ S_n^2\ -\ k^-),\ where\ C_0^-\ =\ 0\]其結果呈現如下
```{r warning=FALSE}
library(knitr)
data = read.table("ex35.dat.txt", header = T)
data$C_n[1] = min(0, 0+data$s[1]^2-k_minus)
for (i in 2:10) {
    data$C_n[i] = min(0, data$C_n[i-1]+data$s[i]^2-k_minus)
}

kable(data.frame(C_n = round(data$C_n, 4)), row.names = 1:10)
```
繪製 CUSUM chart 

```{r}
ii = seq(1,10)
par(mfrow = c(1,1), mar=c(4,4,1,6))
plot(ii,data$C_n,type="o",lty=1,pch=16,xlab="n",
     ylab=expression(C[n]^{"-"}),mgp=c(2,1,0),xlim=c(1,10), 
     ylim=c(-0.8,0),cex=0.8)
lines(ii,rep(h_l,10),lty=2,cex=0.8)
axis(4, at = h_l, labels = "h_l = -0.681", cex=0.8,las=1)
title("CUSUM chart (DI form)")
```
In the 10 samples from a process producing bearings, the CUSUM chart does not give any signals. Therefore, we can not detect downward variance shifts from the observed data. 




\textbf{4.15}  
\[
\begin{aligned}
&\because\ \frac{1}{ARL_{0,J}}\ \approx\ \frac{1}{ARL_{0,M}}\ +\ \frac{1}{ARL_{0,V}},\ and\ ARL_{0,J}\ =\ 200\\
&\Rightarrow\ Make\ \ that\ \ ARL_{0,M}\ =\ ARL_{0,V}\ =\ 400\\
&And\ \ (k,h)\ =\ (0.5,0.128),\ \ (k^+,h_U)\ =\ (0.4621,0.633)\\
&(Note\ \ that\ \ k\ =\ \frac{\mu_1-\mu_2}{2}\ =\ 0.5\ ,\ \\
&and\ \ h\ =\ 0.128\ \ by\ the\ software\ package\ ANYGETH.EXE\\
&k^+\ =\ \frac{2\sigma_0\sigma_1log(\sigma_1/\sigma_0)}{\sigma_1^2-\sigma_0^2}\ =\ 0.4621\ ,\ \\&and\ \ h_U\ =\ 0.633\ by\ the\ software\ package\ ANYGETH.EXE)
\end{aligned}
\]

```{r}
k_mean = 0.5
h_mean = 0.128

k_vplus = (2*0.25*1*log(1/0.5))/(1-0.25)
h_vplus = 0.633
```

匯入資料並計算
$$
\left\{
\begin{aligned}
&CUSUM-M:C_n^+\ =\ max(0\ ,\ C_{n-1}^++(\overline{X}_n-\mu_0)-k)\ ,\ where\ \ C_0^+\ =\ 0\\
&CUSUM-V:C_n^+\ =\ max(0\ ,\ C_{n-1}^++S_n^2-k^+)\ ,\ where\ \ C_0^+\ =\ 0
\end{aligned}
\right.
$$
其結果呈現如下
```{r warning=FALSE}
data415 = read.table("ex415.dat.txt", header = T)
data415$C_mean[1] = max(0, 0+(data415$xbar[1]-1)-k_mean)
for (i in 2:24) {
    data415$C_mean[i] = max(0, data415$C_mean[i-1]+data415$xbar[i]-1-k_mean)
}

data415$C_var[1] = max(0,0+data415$xstd[1]^2-k_vplus)
for (i in 2:24) {
    data415$C_var[i] = max(0, data415$C_var[i-1]+data415$xstd[i]^2-k_vplus)
}
C = data.frame(C_n_M = data415$C_mean, 
               C_n_V = data415$C_var)
kable(round(C, 4), row.names = 1:24)
```

繪製 CUSUM-M 和 CUSUM-V chart
```{r}
par(mfcol = c(1,2), mar=c(5,3.5,4,5.5))
ii = seq(1,24)
plot(ii,data415$C_mean,type="o",lty=1,pch=16,xlab="n",
     ylab=expression(C[n]^{"+"}),mgp=c(2,1,0),xlim=c(1,24), 
     ylim=c(0,6),cex=0.8)
lines(ii,rep(h_mean,24),lty=2,cex=0.8)
axis(4, at=h_mean,labels="h = 0.128",cex=0.8,las=1)
title("CUSUM-M chart")


plot(ii, data415$C_var, type="o",lty=1,pch=16,xlab="n",
     ylab=expression(C[n]^{"+"}),mgp=c(2,1,0),xlim=c(1,24), 
     ylim=c(0,10),cex=0.8)
lines(ii, rep(h_vplus, 24), lty=2,cex=0.8)
axis(4, at = h_vplus, labels = "h_u = 0.633", cex=0.8,las=1)
title("CUSUM-V chart")
```
It can be seen that the CUSUM-M chart gives a signal at the 13th time point, and the CUSUM-V chart also gives a signal at the 13th time point.

```{r}
```


```{r}
```


```{r}
```


```{r}
```

