---
title: "品質管制 Homework 4"
header-includes:
  - \usepackage{xeCJK}
  - \setCJKmainfont{標楷體}
output: 
  pdf_document:
    latex_engine: xelatex
    number_sections: true
author: "110024516 統研碩一 邱繼賢"
date: "2021 年 10 月 21 日"
---

\textbf{3.19}
\textbf{(i)}  

```{r echo=FALSE, message=FALSE, error=FALSE}
df = read.table("ex35.dat.txt", header = T)
data = data.matrix(df[, 1:5])
datac = c(data)
xbar = mean(datac)
s = sd(datac)

qqnorm(datac)
qqline(datac)

```
The data are plotted against a theoretical normal distribution in such a way that the points should form an approximate straight line. Departures from this straight line indicates from normality. Thus, the combined sample has been checked for the normality. 




\textbf{(ii)}  
$\overline{X}\ =\ \frac{1}{50}\sum_{i=1}^{50}{X_i}\ =\ 35.3046$, and $S\ =\ \sqrt{\frac{1}{50-1}\sum_{i=1}^{50}{(X_i\ -\ \bar{X})^2}}\ \approx\ 0.9099$

$P(LSL\ \leq\ X\ \leq\ USL)\ =\ P(\frac{LSL-\overline{X}}{S}\ \leq\ Z\ \leq\ \frac{USL-\overline{X}}{S})\\=\ \Phi(\frac{USL-\overline{X}}{S})\ -\ \Phi(\frac{LSL-\overline{X}}{S})\ \approx\ \Phi(2.9625)\ -\ \Phi(-2.5329)\ \approx\ 0.9928$
```{r include=FALSE}
pnorm((38-xbar)/s) - pnorm((33-xbar)/s)

```


\textbf{(iii)}  
In Exercise 3.5 (i), $\overline{X}\ =\ 35.304,\ and \ \ \hat{\sigma}\ = \ \frac{\overline{R}}{d_1(5)}\ \approx\ 0.8650$

$P(LSL\ \leq\ X\ \leq\ USL)\ =\ P(\frac{LSL-\overline{X}}{\hat{\sigma}}\ \leq\ Z\ \leq\ \frac{USL-\overline{X}}{\hat{\sigma}})\ \approx\ \Phi(3.1167)\ -\ \Phi(-2.6636)\ \approx\ 0.9952$


```{r include=FALSE}
xbar_2 = 35.304
s_2 = 2.012/(2.326)
pnorm((38-xbar_2)/s_2) - pnorm((33-xbar_2)/s_2)
```



\textbf{3.20}
\textbf{(i)}  
$LSL\ =\ 2100, \ USL\ =\ 2300, \ T\ =\ \frac{USL\ +\ LSL}{2}, \ m\ =\ 50, \ \overline{X}\ =\ 2250, \ s\ =\ 50\\\hat{C}_p\ =\ \frac{USL\ -\ LSL}{6s}\ \approx\ 0.6667 \ \\\hat{C}_{pl}\ =\ \frac{\overline{X}\ -\ LSL}{3s}\ =\ 1, \ \hat{C}_{pu}\ =\ \frac{USL\ -\ \overline{X}}{3s}\ \approx\ 0.3333\ \Rightarrow\ \hat{C}_{pk}\ =\ min(\hat{C}_{pl}, \ \hat{C}_{pu})\ =\ 0.3333\\ \hat{C}_{pm}\ =\ \frac{\hat{C}_p}{\sqrt{1+(\frac{\overline{X}\ -\ T}{s})^2}}\ \approx\ 0.4714$  
Between $C_p\ and \ C_{pk}$, $C_{pk}$ is more appropriate to use in this case, because $\overline{X}\ \neq\ T$.   

It is natural to observe that $P(LSL\ <\ X\ <\ USL\ \ |\ \ \overline{X}\ =\ T)\ >\ P(LSL\ <\ X\ <\ USL\ \ |\ \ \overline{X}\ \neq\ T)$, so we prefer to use $C_{pk}$ in this case.



```{r include=FALSE}
LSL = 2100
USL = 2300
T = (LSL + USL)/2
d = (USL - LSL)/2
m = 50
xbar = 2250
s = 50

cp_hat = (USL - LSL)/(6*s)
cpl_hat = (xbar - LSL)/(3*s)
cpu_hat = (USL - xbar)/(3*s)
cpk_hat = min(cpl_hat, cpu_hat)
cpm_hat = cp_hat / sqrt(1+((xbar - T)/s)^2)
```




\textbf{(ii)}  
A 95% confidence interval for $C_p$ is $$(\hat{C}_p\sqrt{\frac{\chi^2_{\frac{\alpha}{2}, m-1}}{m-1}},\ \hat{C}_p\sqrt{\frac{\chi^2_{1-\frac{\alpha}{2}, m-1}}{m-1}})\ \approx\ (0.5350, \ 0.7981)$$

```{r include=FALSE}
lower_bd = cp_hat * sqrt(qchisq(0.025, m-1)/(m-1))
upper_bd = cp_hat * sqrt(qchisq(0.975, m-1)/(m-1))
c(lower_bd, upper_bd)
```




\textbf{(iii)}  
$P(X\ <\ LSL\ or \ X\ >\ USL\ \ |\ \ process\ IC)\ \\=\ P(Z\ <\ \frac{LSL\ -\ \overline{X}}{s})\ +\ P(Z\ >\ \frac{USL\ -\ \overline{X}}{s})\ =\ \Phi(-3)\ +\ (1\ -\ \Phi(1))\ \approx\ 0.1600$
```{r include=FALSE}
pnorm((LSL-xbar)/s) + 1 - pnorm((USL - xbar)/s)

```



\textbf{4.2}
\textbf{(i)}  
$k\ =\ 0.5, \ h\ = \ 3.502, \ \mu_0\ =\ 0$  
The charting statistic $C_n\ =\ \sum_{i=1}^{20}{(X_i\ -\ \mu_0)}$

```{r echo=FALSE, message=FALSE, error=FALSE}
data = read.csv("data4-2.csv", header = T)
mu_0 = 0
k = 0.5
h = 3.502
sum = 0
for (i in 1:20) {
  sum = sum + data$X_i[i]
  data$C_i[i] = sum
}

ii = seq(1,20)
mask1 = rep(0, 20)
mask2 = rep(0, 20)
plot(ii, data$C_i, type="o",lty=1,pch=16,xlab="n",
     ylab=expression(C[n]),mgp=c(2,1,0),xlim=c(0,21), 
     ylim=c(-3,30),cex=0.8)

for (i in 1:20) {
  mask1[i] = data$C_i[20] - h + (i - 20)*0.5
  mask2[i] = data$C_i[20] + h + (20 - i)*0.5
}

lines(ii,mask1,type="l",lty=1,cex=2.5)
lines(ii,mask2,type="l",lty=2,cex=2.5)
lines(rep(20,2),c(data$C_i[20]-h,data$C_i[20]+h),lty=3,cex=2.5)
lines(rep(1,2),c(mask1[1],mask2[1]),lty=3,cex=2.5)

cord.x <- c(1,1,20,20)
cord.y <- c(mask1[1],mask2[1],data$C_i[20]+h,data$C_i[20]-h)
polygon(cord.x,cord.y,density=7,angle=135,border=NA,lty=1,cex=0.6)

title("CUSUM chart", cex = 0.9)
```
The CUSUM chart detects a positive mean shift occurs before n = 20, because there are some values of the charting statistics falling below that half-line.

Then, we choose the time point, which is farthest away from the V-mask, is used as the estimate of $\tau$$$\Rightarrow\ \hat{\tau}\ =\ 10$$

After the estimate of $\tau$, $\hat{\tau}$ is obtained, $\delta$ can be estimated by $$\Rightarrow\ \hat{\delta}\ =\ \frac{C_{20}\ -\ C_{\hat{\tau}}}{20\ -\ \hat{\tau}}\ =\ \frac{14.34\ -\ (-1.81)}{20\ -\ 10}\ =\ 1.615$$



\textbf{(ii)}  
$C_n^+\ =\ max(0,\ C_{n-1}^+\ +\ (X_n\ -\ \mu_0)\ -\ k)\ ,\ \ where\ \ C_0^+\ =\ 0$

```{r echo=FALSE, message=FALSE, error=FALSE}
data$C_plus[1] = max(0, 0+data$X_i[1] - k )
for (i in 2:20) {
  data$C_plus[i] = max(0, data$C_plus[i-1]+data$X_i[i]-k)
}


plot(ii, data$C_plus,type="o",lty=1,pch=16,xlab="n",
     ylab=expression(C[n]^{"+"}),mgp=c(2,1,0),xlim=c(0,21), 
     ylim=c(0,13),cex=0.8)
lines(ii,rep(h,20),lty=2,cex=0.8)
title("DI form CUSUM chart", cex = 0.9)
```
This kind of CUSUM chart also gives a signal of an upward mean shift because $$C_n^+\ >\ h\ =\ 3.502,\ for \ n\ \geq\ 13$$ It shows the same results in part(i). 

```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```

