---
title: "Linear Model Assignment 5"
header-includes:
  - \usepackage{xeCJK}
  - \setCJKmainfont{標楷體}
output: 
  pdf_document:
    latex_engine: xelatex
    number_sections: true
author: "110024516 統研碩一 邱繼賢"
---
\textbf{Problem 1.}  
\textbf{a.}  
```{r}
travel = read.table("travel.txt", skip = 1)
names(travel) = c("obs", "n", "x", "y")
plot(travel$x, travel$y, xlab = "x", ylab = "y", cex = 2)
```


(1) x 和 y 存在著正相關的現象。

(2) x 對 y 普遍存在著低估的現象，即對兩地點移動所需時間估計普遍小於實際測量後的平均值。



\textbf{b.}
 以每組地點間的 travelers 數量為權重($w_i\ \propto\ n_i$)，建構回歸模型如下：
\[S^{-1}Y\ =\ S^{-1}X\beta\ +\ S^{-1}\epsilon\ ,\ \text{where}\ S\ =\ diag(\frac{1}{\sqrt{w_1}},...,\frac{1}{\sqrt{w_n}})\ ,\ \text{then}\ \Sigma\ =\ SS^T\]
```{r}
w = travel$n
g = lm(y ~ x, data = travel, weights = w)
summary(g)
```

```{r echo=FALSE}
plot(travel$x, travel$y, xlab = "x", ylab = "y", cex = 2)
abline(g)
```


\newpage
\textbf{c.}
 In order to check model g for lack of fit. Construct saturated model ga and do the anova test to compare the two models as below :
\[
\left\{
\begin{aligned}
&H_0:\text{g model fitted better}\\
&H_1:\text{ga model fitted better}
\end{aligned}
\right.\ \Longleftrightarrow\ 
\left\{
\begin{aligned}
&H_0:\text{g model is not lack of fit}\\
&H_1:\text{g model is lack of fit}
\end{aligned}
\right.
\]
```{r}
ga = lm(y ~ factor(x), data = travel, weights = w)
anova(g, ga)
```
p-value = $0.07132\ >\ 0.05\ \Rightarrow$ fail to reject $H_0$

$\therefore$ We do not detect lack of fit for model g.




\newpage
\textbf{Problem 2.}  
\textbf{a.}
 Take the number of fathers in each category as weight($w_i\ \propto\ n_i$), then contruct the Weighted Least Square as below :
\[\text{model}\ g1\ :\ S^{-1}Y\ =\ S^{-1}X\beta\ +\ S^{-1}\epsilon\ ,\ \text{where}\ S^{-1}\ =\ diag(\sqrt{w_1},...,\sqrt{w_n})\]

```{r}
height = read.table("height.txt", skip = 2)
father_h = height[,1]
son_h = height[,2]
w = height[,3]
g1 = lm(son_h ~ father_h, weights = w)
summary(g1)
```





\textbf{b.}
 Construct model g2 : $height\ of\ son\ =\ height\ of\ father\ +\ error\ ,\ with\ w_i\ \propto\ n_i$

and then do the anova test for comparing g1 and g2 models : 
\[
\left\{
\begin{aligned}
&H_0:\text{g2 model fits better}\\
&H_1:\text{g1 model fits better}
\end{aligned}
\right.
\]
```{r}
g2 = lm(son_h ~ offset(father_h)-1, weights = w)
anova(g2, g1)
```

$\because\ p-value\ <\ 0.05\ \Rightarrow\ \text{reject}\ H_0$

$\therefore$ g1 model is a better model for fitting. g1 model is not approriate to be simplified to g2.


\newpage
\textbf{Problem 3.}  
(i) 整理 data：  
將 data 的第二欄數據$\times0.00001+0.742$伸縮平移到其對應的真實數據值，並以相同的 day 為一組計算其 standard deviation (std)，然後取 standard variance 為其權重 ($w_i\ \propto\ \frac{1}{std_i^2}$)，整理後資料呈現如下：

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(knitr)
crank = read.table("crank.txt", skip = 1)
names(crank) = c("day", "diameter")
crank$diameter = 0.742+0.00001*crank$diameter

crank = crank %>% group_by(day) %>% 
    mutate(std = sd(diameter)) %>% 
    mutate(weight = 1/std^2) %>% 
    ungroup()
kable(crank)
```


(ii) Test for under control or not  
建構模型
\[
\left\{
\begin{aligned}
&g_3:diameter\ =\ \beta_0\ +\ \beta_1\ day\ +\ \epsilon\ ,\ with\ weight\ \propto\ \frac{1}{std^2}\\
&g_4:diameter\ =\ 0.74275\ +\ \epsilon\ ,\ with\ weight\ \propto\ \frac{1}{std^2}
\end{aligned}
\right.
\]
判斷 process 是否 under control 的條件即為進行以下檢定：
\[
\left\{
\begin{aligned}
&H_0:\beta_0\ =\ 0.74275\ \ and\ \ \beta_1\ =\ 0\\
&H_1:\beta_0\ \neq\ 0.74275\ \ or\ \ \beta_1\ \neq\ 0
\end{aligned}
\right.\ \Leftrightarrow\ 
\left\{
\begin{aligned}
&H_0:g4\ \ fits\ \ better\\
&H_1:g3\ \ fits\ \ better
\end{aligned}
\right.
\]

```{r}
g3 = lm(diameter ~ day, weights = weight, data = crank)
g4 = lm(diameter ~ offset(rep(0.74275,40))-1, 
        weights = weight, data = crank)
anova(g4,g3)
```
$p-value\ <\ 0.05\ \Rightarrow$ right $H_0$  
$\therefore\ g_3$ fits better. The process is out of control. 


(iii) Test for lack of fit  
建構 saturated model $g_5:diameter\ \sim\ factor(day)\ ,\ with\ weight\ \propto\ \frac{1}{std^2}$  
並進行以下檢定
\[
\left\{
\begin{aligned}
&H_0:g_3\ fits\ better\\
&H_1:g_5\ fits\ better
\end{aligned}
\right.\ \Leftrightarrow\ 
\left\{
\begin{aligned}
&H_0:g_3\ is\ not\ lack\ of\ fit\\
&H_1:g_3\ is\ lack\ of\ fit
\end{aligned}
\right.
\]

```{r}
g5 = lm(diameter ~ factor(day), weights = weight, data = crank)
anova(g3, g5)
```
$p-value\ =\ 0.248\ >\ 0.05\ \Rightarrow$ fail to reject $H_0$  
$\therefore$ We do not detect lack of fit for model $g_3$




```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```

