---
title: "Experimental Design and Analysis Homework 6"
author: "110024516 邱繼賢"
header-includes:
  - \usepackage{xeCJK}
  - \setCJKmainfont{標楷體}
  - \linespread{1.5}
output: 
  pdf_document:
    latex_engine: xelatex
---

## Problem 1. (5-4)  
如果一個 design 的 $\text{resolution} = R$，代表在它的 defining contrast subgroup 中最短的 $\text{wordlength} = R$，那在它投影到一個 $R-1$ 的 factorial design 時，由於只有 $R-1$ 個 factors，所以它一定無法形成長度為 $R$ 的 word，故一定為一個 full factorial design。  






## Problem 2. (5-13)  
### (a)  
There are five 2-level factors :  

(1) **temperature** : 160°F (-) or 180°F (+)  
(2) **concentration** : 30% (-) or 40% (+)  
(3) **catalyst** : A (-) or B (+)  
(4) **stirring rate** : 60rpm (-) or 100rpm (+)  
(5) **pH** : low (-) or high (+)  

and we need to avoid the combinations of (+,+, ,+, ) & (+, ,+, ,+), so we construct the $2^{5-2}$ design with 2 generators :
$$
4\ =-12\ ,\ 5\ =-13
$$
then the defining contrast subgroup : 
$$
I\ =-124\ =-135\ =\ 2345
$$
The design matrix as below : 


| **Run** | **temperature** | **concentration** | **catalyst** | **stirring rate** | **pH** |
|:----------------:|:---------------:|:-----------------:|:------------:|:-----------------:|:------:|
| 1                | +               | +                 | +            | -                 | -      |
| 2                | +               | +                 | -            | -                 | +      |
| 3                | +               | -                 | +            | +                 | -      |
| 4                | +               | -                 | -            | +                 | +      |
| 5                | -               | +                 | +            | +                 | +      |
| 6                | -               | +                 | -            | +                 | -      |
| 7                | -               | -                 | +            | -                 | +      |
| 8                | -               | -                 | -            | -                 | -      |

此處僅為展示方便，實際做實驗時需要將實驗順序進行隨機排序。  


### (b)  
Use defining contrast subgroup : 
$$
I\ =-124\ =\ 345\ =-1235
$$
and there are seven alias sets : 
$$
\begin{aligned}
&1=-24=1345=-235\\
&2=-14=2345=-135\\
&3=-1234=45=-125\\
&12=-4=12345=-35\\
&13=-234=145=-25\\
&23=-134=245=-15\\
&123=-34=1245=-5
\end{aligned}
$$
We can see that the catalyst-by-temperature (13) and catalyst-by-concentration (23) interaction effects are neither aliased with the main effects nor with each other. The design matrix is shown below : 

| **Run** | **temperature** | **concentration** | **catalyst** | **stirring rate** | **pH** |
|:-------:|:---------------:|:-----------------:|:------------:|:-----------------:|:------:|
| 1       | +               | +                 | +            | -                 | -      |
| 2       | +               | +                 | -            | -                 | +      |
| 3       | +               | -                 | +            | +                 | +      |
| 4       | +               | -                 | -            | +                 | -      |
| 5       | -               | +                 | +            | +                 | +      |
| 6       | -               | +                 | -            | +                 | -      |
| 7       | -               | -                 | +            | -                 | -      |
| 8       | -               | -                 | -            | -                 | +      |

我們為了在此實驗設計中避免某些 2-factor interactions 和主效應有所混淆，相對的就無法避免 (+, ,+, ,+) 這種高風險組合的出現 (look at the 3rd run)。  
一樣此處僅為展示方便，實際進行實驗仍須將實驗順序隨機排列。  



## Problem 3. (5-15)  
### (a)  
Let's look at the defining contrast subgroups of both two designs : 
$$
\begin{aligned}
&\text{(i)}\ I\ =\ 12345\ =\ 1246\ =\ 356\ \Rightarrow\ \text{resolution}\ =\ 3\\
&\text{(ii)}\ I\ =\ 1235\ =\ 1246\ =\ 3456\ \Rightarrow\ \text{resolution}\ =\ 4
\end{aligned}
$$
I prefer design **(ii)** because it has larger resolution.  

### (b)  
Take a look at only the alias sets, which contain 2-factor interaction effects, of design **(ii)** : 
$$
\begin{aligned}
&12=35=46=123456\\
&13=25=2346=1456\\
&14=2345=26=1356\\
&23=15=1346=2456\\
&24=1345=16=2356\\
&34=1245=1236=56\\
&1234=45=36=1256
\end{aligned}
$$
We can ignore all the 2-factor interactions involving factor 6 and 3(or higher)-factor interactions in above alias sets.  
Therefore, 2-factor interactions : **14, 24, 34, 45** are estimable.  
However, **12=35, 13=25, 23=15** these three pairs of 2-factor interactions are still aliased.  


### (c)  
Arranging the $2^{6-2}$ design **(ii)** in $2^1$ blocks need 1 block factor. My blocking scheme is $B\ =\ 134$ because the confounding set under this condition is 
$$
B\ =\ 134\ =\ 245\ =\ 236\ =\ 156
$$
only contains 3-factor interactions which are negligible under the usual assumption in **(b)**. In this way, we will not confound any other main or 2-factor interaction effects which we more concern about.  


## Problem 4. (5-16)  
### (a)  
Let's look at the defining contrast subgroups of two $2^{6-2}$ designs : 
$$
\begin{aligned}
&\text{A}\ :\ I\ =\ 12345\ =\ 1236\ =\ 456\ \Rightarrow\ \text{resolution}\ =\ 3\\
&\text{B}\ :\ I\ =\ 1235\ =\ 2346\ =\ 1456\ \Rightarrow\ \text{resolution}\ =\ 4
\end{aligned}
$$
I will choose design B because it has larger resolution.  

### (b)  
There are $2^{4}=16$ runs in $2^{6-2}$ design, so we only have 15 degrees of freedom to estimate factorial effects. However, if resolution = 5, all main and 2-factor interaction effects are clear, namely all of them are fall in different alias sets. We will need at least $6+\begin{pmatrix}6\\2 \end{pmatrix}\ =\ 21$ degrees of freedom to estimate all of them. It is impossible.  



## Problem 5. (5-28)  
### (a)  
All 2-factor interaction effects : **AB, AC, AD, AE, BC, BD, BE, CD, CE, DE** are clear.  
All main effects : **A, B, C, D, E** are strongly clear.  


### (b)  
計算每一組 level combinations 下的 $\bar{y}$ 和 $s^2$，若 $s^2=0$ 則以 $0.001$ 代入

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(knitr)
welding = read.table("welding.txt", header = T)
colnames(welding) = c("A","B","C","D","E","y")
data = welding %>% group_by(A,B,C,D,E) %>% 
    summarise(y_bar = mean(y), s_square = ifelse(var(y)==0,0.001,var(y))) %>% 
    ungroup() %>% 
    mutate(A=rep(c(-1,1),each=8),B=rep(c(-1,1),each=4,2),C=rep(c(-1,1),each=2,4),D=rep(c(-1,1),8), 
           E=c(-1,1,1,-1,1,-1,-1,1,1,-1,-1,1,-1,1,1,-1))
kable(data, col.names = c("A","B","C","D","E","$\\bar{y}$","s^2"), digits = 3)
```


根據 **(a)** 所得到的結論，可以得知每一個 main 和 2-factor interaction effects 都是 clear effects，也就是說它們都只和 3(or 4)-factor interactions 落在同一個 alias sets，而那些高階的 interaction effects 都是可以忽略它們的效應的，故我們可以建構以下 location 和 dispersion models : 
$$
\bar{y}\ =\ X\beta\ +\ \epsilon\ \ ,\ \ \ln s^2\ =\ X\gamma\ +\ \delta
$$
where $X$ contains all main and 2-factor interaction effects of the five factors.  
然後可以藉由估計出來的迴歸係數來推得 location 和 dispersion effects :
$$
\hat{\theta}\ =\ 2\hat{\beta}\ \ ,\ \ \hat{\psi}\ =\ 2\hat{\gamma}
$$
結果呈現如下表：

```{r}
loc_mod = lm(y_bar ~ A+B+C+D+E+A:B+A:C+A:D+A:E+B:C+B:D+B:E+C:D+C:E+D:E, data)
loc_effect = 2*coef(loc_mod)[-1]
dis_mod = lm(log(s_square) ~ A+B+C+D+E+A:B+A:C+A:D+A:E+B:C+B:D+B:E+C:D+C:E+D:E, data)
dis_effect = 2*coef(dis_mod)[-1]
effect_tab = data.frame(y_bar = loc_effect, log_s = dis_effect)
kable(effect_tab, col.names = c("$\\bar{y}$","$ln s^2$"), digits = 3)
```

接下來以 Half-Normal plot 的方式來判斷哪些效應為顯著：

```{r}
"halfnorm" <- function (x, nlab = 2, labs = as.character(1:length(x)), ylab = "Sorted Data") {
    x <- abs(x)
    labord <- order(x)
    x <- sort(x)
    i <- order(x)
    n <- length(x)
    ui <- qnorm((n + 1:n)/(2 * n + 1))
    plot(ui, x[i], xlab = "Half-normal quantiles", ylab = ylab, ylim=c(0,max(x)),
         type = "n")
    if(nlab < n)
        points(ui[1:(n - nlab)], x[i][1:(n - nlab)])
    text(ui[(n - nlab + 1):n], x[i][(n - nlab + 1):n], labs[labord][(n - nlab + 1):n])
}
par(mfrow = c(1,2))
halfnorm(loc_effect,3,names(loc_effect), "Absolute Effect")
title("Location Effect Half-Normal plot")
halfnorm(dis_effect,3,names(dis_effect), "Absolute Effect")
title("Dispersion Effect Half-Normal plot")
```

從圖形判斷：  

(1) Location : A, AC 兩效應顯著  
(2) Dispersion : C, D, CD 三效應顯著  

故能配飾 location 和 dispersion models :  
$$
\begin{aligned}
&\hat{\bar{y}}\ =\ \hat{\beta}_0\ +\ \hat{\beta}_A\ x_A\ +\ \hat{\beta}_{AC}\ x_{AC}\ =\ 1605\ +\ 263.5\ x_A\ +\ 82.5\ x_{AC}\\
&\ln\hat{s}^2\ =\ \hat{\gamma}_0\ +\ \hat{\gamma}_C\ x_C\ +\ \hat{\gamma}_D\ x_D\ +\ \hat{\gamma}_{CD}\ x_{CD}\ =\ 5.3456\ -\ 2.715\ x_C\ +\ 2.5463\ x_D\ +\ 2.5003\ x_{CD}
\end{aligned}
$$

### (c)  

```{r message=FALSE, warning=FALSE}
library(latex2exp)
data2 = data %>% group_by(C,D) %>% summarise(m = mean(log(s_square)))
plot(data2$C, data2$m, xaxt = "n",xlab = "C", ylab = TeX("$ln s^2$"), 
     main = "Interaction plot of C : D", xlim = c(-1.5,1.5), pch = 1)
axis(1,c(-1,1),labels = c("-","+"))
segments(-1,data2$m[1],1,data2$m[3]) ; segments(-1,data2$m[2],1,data2$m[4])
text(1,data2$m[3],"D -",pos = 4) ; text(1,data2$m[4],"D +",pos = 4)
```

藉由 factor C, D 的 interaction plot 可以看出在 $(C, D)$ 不同的設定值下，$\ln s^2$ 的大小也會有所不同，而從圖中看出 $(C,D)\ =\ (+,-)$ 會有最小的變異，從 dispersion model 觀察也能得到相同結論，故 $(C,D)=(+,-)$ 就是 optimal factor settings for minimizing variance  

### (d)  
選擇 factor settings $(A,C)\ =\ (+,+)$ 可以得到 location model 的最大值
$$
\hat{\bar{y}}\ =\ 1605\ +\ 263.5\ +\ 82.5\ =\ 1951
$$
故 $(A,C)\ =\ (+,+)$ 就是 optimal factor settings for maximizing the tensile strength  


### (e)  
這是一個 Larger-the-better problem，adjustment factor 為那些對 dispersion model 有顯著效應但是對 location model 無顯著貢獻的 factor，故 factor D 為 adjustment factor  
Two-step procedure :  

(1) Choose $(A,C)\ =\ (+,+)$ to maximize the tensile strength
$$
\hat{\bar{y}}\ =\ 1605\ +\ 263.5\ +\ 82.5\ =\ 1951
$$
(2) Choose $(C,D)=(+,-)$ to minimize variance
$$
\hat{s}^2\ =\ \exp\left[5.3456\ -\ 2.715\ -\ 2.5463\ -\ 2.5003\right]\ =\ 0.08927802
$$

PS : Larger-the-best 和 Nominal-the-best problem 在決定 adjustment factor 以及 two-step procedure 順序剛好相反。










































































