---
title: "Statistical Computing HW1"
author: "110024516 統研碩一 邱繼賢"

output:
  html_document:
    toc: true
    toc_depth: 3
    collapsed: false
    smooth_scroll: false
---

## (1) Weibull distribution  

$$
\begin{aligned}
&X\ \sim\ F(x)\ =\ 1-e^{-(\frac{x}{\theta})^\beta}\ \ ,\ \theta > 0\ ,\ \beta>0\ ,\ x>0\\
&\Rightarrow\ F^{-1}(p)\ =\ \theta[-log(1-p)]^{\frac{1}{\beta}}\\
&\text{Let}\ \ U\ \sim\ U(0,1)\ ,\ \text{and}\ \because\ F(X)\ \stackrel{d}{=}\ U\ \sim\ U(0,1)\\ 
&\therefore\ X\ \stackrel{d}{=}\ F^{-1}(U)\ =\ \theta[-log(1-U)]^{\frac{1}{\beta}}\ \sim\ F\\
\end{aligned}
$$

Therefore, we can draw 1000 samples from the *U(0,1)* distribution, and transform the samples by the inverse cdf to represent the original distribution (namely Weibull distribution).  

Take $(\theta\ ,\ \beta)\ =\ (1,5)$ for example. The results are shown as below : 


```{r echo=FALSE, message=FALSE, warning=FALSE}
library(latex2exp)
# (1)
F1_inverse = function(x, theta, beta) {
    theta*(-log(1-x))^(1/beta)
}
f1 = function(x, theta, beta) {
    (beta*x^(beta-1))*exp(-(x/theta)^beta)/(theta^beta)
}
set.seed(09345)
n = 1000
y = runif(n)
x = F1_inverse(y,1,5)
hist(x, 15,probability = T, ylim = c(0,2), main = "real / empirical pdf")
curve(f1(x,1,5), 0,4, col = 4, lwd = 2, add = T)
legend("topright", legend = TeX("$f_1(x)$"), col = 4, lwd = 2)
box()
```


## (2) Pareto distribution  

$$
\begin{aligned}
&\text{Target pdf : }\ f_2(x\ ;\ \theta,\beta)\ =\ \frac{\beta}{\theta(1+\frac{x}{\theta})^{\beta+1}}\ ,\ \theta>0\ ,\ \beta>0\ ,\ x>0\\
&\text{Proposal pdf : }\ g_2(x\ ;\ \nu)\ =\ 2\ h(x\ ;\ \nu)\ ,\ x>0\ ,\ \nu>0\ ,\\
&\text{where}\ h(x\ ;\ \nu)\ \text{is the pdf of t distribution with degree of freedom}\ =\ \nu\\
&\text{Compute the value}\ M_2\ \text{, such that}\ M_2\ \geq\ Max_{x>0}\frac{f_2(x\ ;\ \theta,\beta)}{g_2(x\ ;\ \nu)}\\
&\text{Generate}\ Y\ \sim\ g_2\ \text{and}\ U\ \sim\ U(0,1)\ \text{.}\\
&\text{Let}\ X=Y\ \text{if}\ U\ \leq\ \frac{f_2(Y)}{M_2\ g_2(Y)}\ \text{, and reject the draw otherwise.}\\
&\text{Hence, the theoretical efficiency :}\ P(U\ \leq\ \frac{f_2(Y)}{M_2\ g_2(Y)})\ =\ \frac{1}{M_2}
\end{aligned}
$$

Take $(\theta\ ,\ \beta\ ,\ \nu)\ =\ (15\ ,\ 10\ ,\ 1)$ for example, and find a sufficient large $M_2\ =\ 1.411$ s.t. the theoretical efficiency $=\ \frac{1}{M_2}\ \approx\ 70.87\%$. The results are shown as below : 

```{r echo=FALSE, message=FALSE, warning=FALSE}
f2 = function(x, theta, beta) {
    beta/(theta*(1+x/theta)^(beta+1))
}
par(mfrow = c(1,1))
curve(f2(x, 15,10), 0,10, ylim = c(0,1), lwd = 2, ylab = "", main = "proposal / target pdf")
curve(1.411*2*dt(x,1), 0,10 ,add = T, col = 2, lwd = 2)
curve(2*dt(x,1), 0,10 ,add = T, col = 2, lty = 2, lwd = 2) # proposal pdf : t_1
legend("topright", legend = c(TeX("$f_2(x)$"), TeX("$g_2(x)$"), TeX("$M_2$\\times$g_2(x)$")), 
       col = c(1,2,2), lwd = 2, lty = c(1,2,1))
```

Draw 1000 samples separately from $g_2$ and $U(0,1)$, and follow the rejection rule as above to construct our target pdf $f_2(x)$. The results are shown as below : 

```{r echo=FALSE, message=FALSE, warning=FALSE}
set.seed(08561)
M2 = 1.411
u2 = runif(1000)
y2 = abs(rt(1000,1))
x2 = y2[u2 <= f2(y2, 15, 10)/(M2*2*dt(y2,1))]
hist(x2, 50, prob = T, ylim = c(0,0.7), xlim = c(0,10), xlab = "x", main = "real / empirical pdf")
curve(f2(x,15,10), 0,10, add = T, col = 4, lwd = 2)
legend("topright", legend = TeX("$f_2(x)$"), col = 4, lwd = 2)
box()
```

Finally, compute the empirical efficiency $=\ 71.5\%$

```{r include=FALSE}
length(x2)/1000
```



## (3) Skewed distribution I  

$$
\begin{aligned}
&\text{Target pdf : }\ f_3(x\ ;\ \gamma)\ =\ \frac{2}{\gamma+\frac{1}{\gamma}}\ \phi(x\gamma^{-sign(x)})\ ,\ x\in R\ ,\ \gamma\ >\ 0\\
&\text{Proposal pdf : }\ g_3(x\ ;\ \mu,\sigma^2)\ \sim\ N(\mu\ ,\ \sigma^2)\ \ ,\ x\ \in\ R\ ,\ \mu\in R\ ,\ \sigma^2>0\\
&\text{Compute the value}\ M_3\ \text{, such that}\ M_3\ \geq\ Max_{x\in R}\frac{f_3(x\ ;\ \gamma)}{g_3(x\ ;\ \mu,\sigma^2)}\\
&\text{Generate}\ Y\ \sim\ g_3\ \text{and}\ U\ \sim\ U(0,1)\ \text{.}\\
&\text{Let}\ X=Y\ \text{if}\ U\ \leq\ \frac{f_3(Y)}{M_3\ g_3(Y)}\ \text{, and reject the draw otherwise.}\\
&\text{Hence, the theoretical efficiency :}\ P(U\ \leq\ \frac{f_3(Y)}{M_3\ g_3(Y)})\ =\ \frac{1}{M_3}
\end{aligned}
$$

Take $(\gamma\ ,\ \mu\ ,\ \sigma^2)\ =\ (1.5\ ,\ 0.3\ ,\ 1.5)$ for example, and find a sufficient large $M_3\ =\ 1.42$ s.t. the theoretical efficiency $=\ \frac{1}{M_3}\ \approx\ 70.42\%$. The results are shown as below : 

```{r echo=FALSE, message=FALSE, warning=FALSE}
f3 = function(x, gamma) {
  2/(gamma+1/gamma)*dnorm(x*gamma^(-sign(x)))
}


curve(f3(x,1.5), -7,7, lwd = 2, ylim = c(0,0.5), ylab = "", main = "proposal / target pdf")
curve(dnorm(x,0.3,1.5), -7,7, lwd = 2, col = 2, add = T) # proposal pdf : N(0.3,1.5)
curve(1.42*dnorm(x,0.3,1.5), -7,7,lwd = 2, col = 2, lty = 2, add = T)
legend("topright", legend = c(TeX("$f_3(x)$"), TeX("$g_3(x)$"), TeX("$M_3$\\times$g_3(x)$")), 
       col = c(1,2,2), lwd = 2, lty = c(1,2,1))
```

Draw 1000 samples separately from $g_3$ and $U(0,1)$, and follow the rejection rule as above to construct our target pdf $f_3(x)$. The results are shown as below :

```{r echo=FALSE, message=FALSE, warning=FALSE}
set.seed(09132)
M3 = 1.42
u3 = runif(1000)
y3 = rnorm(1000, 0.3,1.5)
x3 = y3[u3 <= f3(y3,1.5)/(M3*dnorm(y3,0.3,1.5))]
hist(x3, 25,prob = T, xlim = c(-4,6), ylim = c(0,0.5), main = "real / empirical pdf")
curve(f3(x,1.5), -4,6, add = T, col = 4, lwd = 2)
legend("topright", legend = TeX("$f_3(x)$"), col = 4, lwd = 2)
box()
```

Finally, compute the empirical efficiency $=\ 70.3\%$

```{r include=FALSE}
length(x3)/1000
```



## (4) Skewed distribution II  

### (a)  

$$
\begin{aligned}
&\text{Target pdf : }\ f_{4.1}(x\ ;\ \alpha,\nu_1)\ =\ 2h(x\ ;\ \nu_1)\ G(\alpha x\ ;\ \nu_1)\ ,\ x\in R\ ,\ \alpha\in R\ ,\ \nu_1>0\\
&\text{where}\ h(x)\ \text{and}\ G(x)\ \text{are the pdf and cdf of t distribution with}\ \nu_1\ \text{degree of freedom}\\
&\text{Proposal pdf : }\ g_{4.1}(x\ ;\ \nu_2)\ \sim\ t(\nu_2)\ \ ,\ x\ \in\ R\ ,\ \nu_2>0\\
&\text{Compute the value}\ M_{4.1}\ \text{, such that}\ M_{4.1}\ \geq\ Max_{x\in R}\frac{f_{4.1}(x\ ;\ \alpha,\nu_1)}{g_{4.1}(x\ ;\ \nu_2)}\\
&\text{Generate}\ Y\ \sim\ g_{4.1}\ \text{and}\ U\ \sim\ U(0,1)\ \text{.}\\
&\text{Let}\ X=Y\ \text{if}\ U\ \leq\ \frac{f_{4.1}(Y)}{M_{4.1}\ g_{4.1}(Y)}\ \text{, and reject the draw otherwise.}\\
&\text{Hence, the theoretical efficiency :}\ P(U\ \leq\ \frac{f_{4.1}(Y)}{M_{4.1}\ g_{4.1}(Y)})\ =\ \frac{1}{M_{4.1}}
\end{aligned}
$$

Take $(\alpha\ ,\ \nu_1\ ,\ \nu_2)\ =\ (1\ ,\ 2\ ,\ 1)$ for example, and find a sufficient large $M_{4.1}\ =\ 2.015$ s.t. the theoretical efficiency $=\ \frac{1}{M_{4.1}}\ \approx\ 49.63\%$. The results are shown as below :

```{r echo=FALSE, message=FALSE, warning=FALSE}
f4.1 = function(x,alpha, nu) {
  2*dt(x,nu)*pt(alpha*x, nu)
}


curve(f4.1(x,1,2), -7,7, lwd = 2, ylim = c(0,0.7), ylab = "", main = "proposal / target pdf")
curve(dt(x,1), -7,7,lwd = 2, col = 2, add = T)
curve(2.015*dt(x,1), -7,7,lwd = 2, col = 2, add = T, lty = 2)
legend("topright", legend = c(TeX("$f_{4.1}(x)$"), TeX("$g_{4.1}(x)$"), TeX("$M_{4.1}$\\times$g_{4.1}(x)$")), 
       col = c(1,2,2), lwd = 2, lty = c(1,2,1))
```

Draw 1000 samples separately from $g_{4.1}$ and $U(0,1)$, and follow the rejection rule as above to construct our target pdf $f_{4.1}(x)$. The results are shown as below :

```{r echo=FALSE, message=FALSE, warning=FALSE}
set.seed(09301)
M4.1 = 2.015
u4.1 = runif(1000)
y4.1 = rt(1000,1)
x4.1 = y4.1[u4.1 <= f4.1(y4.1,1,2)/(M4.1*dt(y4.1,1))]
hist(x4.1, 50, xlim = c(-4,7), prob = T, ylim = c(0,0.5), main = "real / empirical pdf")
curve(f4.1(x,1,2), -4,7, lwd = 2, col = 4, add = T)
legend("topright", legend = TeX("$f_{4.1}(x)$"), lwd = 2, col = 4)
box()
```

Finally, compute the empirical efficiency $=\ 48\%$

```{r include=FALSE}
length(x4.1)/1000
```


### (b)  

Now I decided to construct the skewed-double-exponential distribution, then $G(x)$ and $h(x)$ are the cdf and pdf of double-exponential distribution. Later, I will use both methods in Problems (3) and (4).

$$
\begin{aligned}
&\textbf{Method 1.}\\
&h(x\ ;\ \lambda)\ =\ \frac{1}{2}\lambda e^{-\lambda|x|}\ ,\ x\in R\ ,\ \lambda>0\\
&G(x\ ;\ \lambda)\ =\ \int_{-\infty}^x h(u\ ;\ \lambda)\ du\ =\ \frac{1}{2}[1+sign(x)(1-e^{-\lambda|x|})]\ ,\ x\in R\ ,\ \lambda>0\\
&\text{By method in Problem (3), define}\\
&\text{Target pdf :}\ f_{4.2.1}(x\ ;\ \lambda,\gamma)\ =\ \frac{2}{\gamma+\frac{1}{\gamma}}\ h(x\gamma^{-sign(x)}\ ;\ \lambda)\ ,\ x\in R\ ,\ \lambda>0\ ,\ \gamma>0\\
&\text{Proposal pdf : }\ g_{4.2.1}(x\ ;\ \nu)\ \sim\ t(\nu)\ \ ,\ x\ \in\ R\ ,\ \nu>0\\
&\text{Compute the value}\ M_{4.2.1}\ \text{, such that}\ M_{4.2.1}\ \geq\ Max_{x\in R}\frac{f_{4.2.1}(x\ ;\ \lambda,\gamma)}{g_{4.2.1}(x\ ;\ \nu)}\\
&\text{Generate}\ Y\ \sim\ g_{4.2.1}\ \text{and}\ U\ \sim\ U(0,1)\ \text{.}\\
&\text{Let}\ X=Y\ \text{if}\ U\ \leq\ \frac{f_{4.2.1}(Y)}{M_{4.2.1}\ g_{4.2.1}(Y)}\ \text{, and reject the draw otherwise.}\\
&\text{Hence, the theoretical efficiency :}\ P(U\ \leq\ \frac{f_{4.2.1}(Y)}{M_{4.2.1}\ g_{4.2.1}(Y)})\ =\ \frac{1}{M_{4.2.1}}
\end{aligned}
$$

Take $(\lambda\ ,\ \gamma\ ,\ \nu)\ =\ (1\ ,\ 1.5\ ,\ 1)$ for example, and find a sufficient large $M_{4.2.1}\ =\ 1.989$ s.t. the theoretical efficiency $=\ \frac{1}{M_{4.2.1}}\ \approx\ 50.28\%$. The results are shown as below :

```{r echo=FALSE, message=FALSE, warning=FALSE}
h = function(x,lambda) {
  0.5*dexp(abs(x),lambda)
}
G = function(x,lambda) {
  0.5*(1+sign(x)*(1-exp(-lambda*abs(x))))
}

f4.2.1 = function(x,lambda,gamma) {
  2/(gamma+1/gamma)*h(x*gamma^(-sign(x)),lambda)
}
curve(f4.2.1(x,1,1.5), -7,7, lwd = 2, ylim = c(0,0.65), ylab = "", main = "proposal / target pdf")
curve(dt(x,1), -7,7, lwd = 2, col = 2, add = T)
M4.2.1 = 1.989
curve(M4.2.1*dt(x,1), -7,7, lwd = 2, col = 2, lty = 2, add = T)
legend("topright", legend = c(TeX("$f_{4.2.1}(x)$"), TeX("$g_{4.2.1}(x)$"), TeX("$M_{4.2.1}$\\times$g_{4.2.1}(x)$")), 
       col = c(1,2,2), lwd = 2, lty = c(1,1,2))
```

Draw 1000 samples separately from $g_{4.2.1}$ and $U(0,1)$, and follow the rejection rule as above to construct our target pdf $f_{4.2.1}(x)$. The results are shown as below :

```{r echo=FALSE, message=FALSE, warning=FALSE}
set.seed(10132)
u4.2.1 = runif(1000)
y4.2.1 = rt(1000,1)
x4.2.1 = y4.2.1[u4.2.1 <= f4.2.1(y4.2.1,1,1.5)/(M4.2.1*dt(y4.2.1,1))]
hist(x4.2.1, 30, probability = T, xlab = "x", ylim = c(0,0.5), 
     main = "real / empirical pdf")
curve(f4.2.1(x,1,1.5),-7,7, col = 4, lwd = 2, add = T)
legend("topright", legend = TeX("$f_{4.2.1}(x)$"), col = 4, lwd = 2)
box()
```

Finally, compute the empirical efficiency $=\ 47.8\%$

```{r include=FALSE}
length(x4.2.1)/1000
```



$$
\begin{aligned}
&\textbf{Method 2.}\\
&\text{By method in Problem (4), define}\\
&\text{Target pdf :}\ f_{4.2.2}(x\ ;\ \lambda,\alpha)\ =\ 2h(x\ ;\ \lambda)G(\alpha x\ ;\ \lambda)\ ,\ x\in R\ ,\ \lambda>0\ ,\ \alpha\in R\\
&\text{Proposal pdf : }\ g_{4.2.2}(x\ ;\ \nu)\ \sim\ t(\nu)\ \ ,\ x\ \in\ R\ ,\ \nu>0\\
&\text{Compute the value}\ M_{4.2.2}\ \text{, such that}\ M_{4.2.2}\ \geq\ Max_{x\in R}\frac{f_{4.2.2}(x\ ;\ \lambda,\alpha)}{g_{4.2.2}(x\ ;\ \nu)}\\
&\text{Generate}\ Y\ \sim\ g_{4.2.2}\ \text{and}\ U\ \sim\ U(0,1)\ \text{.}\\
&\text{Let}\ X=Y\ \text{if}\ U\ \leq\ \frac{f_{4.2.2}(Y)}{M_{4.2.2}\ g_{4.2.2}(Y)}\ \text{, and reject the draw otherwise.}\\
&\text{Hence, the theoretical efficiency :}\ P(U\ \leq\ \frac{f_{4.2.2}(Y)}{M_{4.2.2}\ g_{4.2.2}(Y)})\ =\ \frac{1}{M_{4.2.2}}
\end{aligned}
$$

Take $(\lambda\ ,\ \alpha\ ,\ \nu)\ =\ (1\ ,\ 0.5\ ,\ 1)$ for example, and find a sufficient large $M_{4.2.2}\ =\ 1.756$ s.t. the theoretical efficiency $=\ \frac{1}{M_{4.2.2}}\ \approx\ 56.95\%$. The results are shown as below :


```{r echo=FALSE, message=FALSE, warning=FALSE}
f4.2.2 = function(x,lambda,alpha) {
  2*h(x,lambda)*G(alpha*x,lambda)
}
M4.2.2 = 1.756
curve(f4.2.2(x,1,0.5), -5,9, lwd = 2, ylim = c(0,0.6), ylab = "", main = "proposal / target pdf")
curve(dt(x,1), -5,9, lwd = 2, col = 2, add = T)
curve(M4.2.2*dt(x,1), -5,9, col = 2, lwd = 2, lty = 2, add = T)
legend("topright", legend = c(TeX("$f_{4.2.2}(x)$"), TeX("$g_{4.2.2}(x)$"), TeX("$M_{4.2.2}$\\times$g_{4.2.2}(x)$")), 
       col = c(1,2,2), lwd = 2, lty = c(1,1,2))
```

Draw 1000 samples separately from $g_{4.2.2}$ and $U(0,1)$, and follow the rejection rule as above to construct our target pdf $f_{4.2.2}(x)$. The results are shown as below :

```{r echo=FALSE}
set.seed(10271)
u4.2.2 = runif(1000)
y4.2.2 = rt(1000, 1)
x4.2.2 = y4.2.2[u4.2.2 <= f4.2.2(y4.2.2,1,0.5)/(M4.2.2*dt(y4.2.2,1))]
hist(x4.2.2, 30, prob = T, xlab = "x", 
     main = "real / empirical pdf", ylim = c(0,0.5))
curve(f4.2.2(x,1,0.5),-4,7, col = 4, lwd = 2, add = T)
legend("topright", legend = TeX("$f_{4.2.2}(x)$"), col = 4, lwd = 2)
box()
```

Finally, compute the empirical efficiency $=\ 59.3\%$

```{r include=FALSE}
length(x4.2.2)/1000
```



## (5) 2-dimension distribution  

The target pdf : 
$$
f(x,y)\ =\ 2(1-x)(1-y)(1-xy)^{-3}\ ,\ 0<x<1\ ,\ 0<y<1
$$

```{r echo=FALSE, message=FALSE, warning=FALSE}
f5 = function(x) {
    2*(1-x[1])*(1-x[2])*(1-x[1]*x[2])^(-3)
}
grid = seq(0,1, length = 30)
grid2 = seq(0,0.999, length = 30)
xy = expand.grid(x = grid, y = grid)
xy2 = expand.grid(x = grid2, y = grid)
z = apply(xy, 1, f5)
z2 = apply(xy2, 1, f5)


par(mfrow = c(1,1))
persp(grid, grid, matrix(z,30,30), theta = 30, 
      zlab = "f(x,y)", xlab = "x", ylab = "y")

par(mfrow = c(1,2))
contour(grid, grid, matrix(z,30,30), 
        levels= round(quantile(z2, 0.1*(1:9)),2), 
        xlab = "x", ylab = "y")

image(grid,grid, matrix(z,30,30), xlab = "x", ylab = "y")
```

我們可以發現 *f(x,y)* 有以下特徵：  
1. 圖形對稱於 *y=x*  
2. *(x,y)* 皆偏小和皆偏大處 *pdf* 數值較高，特別是偏大處(即圖形右上角)

Let 
$$
\left\{
\begin{aligned}
u\ =\ log\frac{x}{1-x}\\
v\ =\ log\frac{y}{1-y}
\end{aligned}
\right.\ 
\Rightarrow\ 
\left\{
\begin{aligned}
x\ =\ \frac{e^u}{1+e^u}\\
y\ =\ \frac{e^v}{1+e^v}
\end{aligned}
\right.
$$
And then transform the target pdf into
$$
g(u,v)\ =\ f(\frac{e^u}{1+e^u}\ ,\ \frac{e^v}{1+e^v})|J|\ =\ \frac{2e^ue^v}{(1+e^u+e^v)^3}\ ,\ u\in R\ ,\ v\in R
$$
Compute the marginal pdf, marginal cdf and joint cdf of *g(u,v)*
$$
\left\{
\begin{aligned}
&g_1(u)\ =\ e^u(1+e^u)^{-2}\ \ and\ \ G_1(u)\ =\ 1-(1+e^u)^{-1}\ ,\ u\in R\\
&g_2(v)\ =\ e^v(1+e^v)^{-2}\ \ and\ \ G_2(v)\ =\ 1-(1+e^v)^{-1}\ ,\ v\in R\\
&G(u,v)\ =\ (1+e^u+e^v)^{-1}\ \ ,\ \ u\in R\ ,\ v\in R
\end{aligned}
\right.
$$
$$
\Rightarrow\ 
\left\{
\begin{aligned}
G_1(U)\ \overset{d}{=}\ A\ \sim\ U(0,1)\\
G_2(V)\ \overset{d}{=}\ B\ \sim\ U(0,1)
\end{aligned}
\right.
$$
By *Sklar's Theorem* :
$$
\begin{aligned}
G(u,v)\ &=\ P(U\leq u\ ,\ V\leq v)\ =\ P(G_1(U)\leq G_1(u)\ ,\ G_2(V)\leq G_2(v))\ =\ P(G_1(U)\leq a\ ,\ G_2(V)\leq b)\\
&=\ C(a\ ,\ b)\ =\ (1+\frac{a}{1-a}+\frac{b}{1-b})^{-1}\ ,\ 0<a<1\ ,\ 0<b<1
\end{aligned}
$$
And
$$
c(a\ ,\ b)\ =\ \frac{\partial^2C}{\partial a\partial b}\ =\ 2(1-a)(1-b)(1-ab)^{-3}\ (\text{The same form as }f(x,y))
$$
which is the joint pdf of *a* and *b*, where $A\ \sim\ U(0,1)$ and $B|A=a^*\ \sim\ f_{B|A=a^*}(b)\ =\ 2(1-a^*)(1-b)(1-a^*b)^{-3\ ,\ 0<b<1}$

進行以下步驟即可抽出 *c(a,b)* (即*f(x,y)*)的 empirical pdf  
1. 抽取 10000 個樣本 $a_i\ \sim\ U(0,1)\ ,\ i\ =\ 1,2,...,10000$  
2. 對每個 $a_i$ 建構 conditional distribution $f_{B|A=a_i}$視為 target distribuion，並以*U(0,1)*為 proposal distribution 進行 rejection sampling (一個$a_i$僅抽出一個相對應的$b_i$)  
3. $(a_i,b_i)\ ,\ i=1,2,...,10000$ 即為從 *c(a,b)(f(x,y))*所抽出的樣本






```{r echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
copula_pdf = function(a,b) {
    2*(1-a)*(1-b)*(1-a*b)^(-3)
}
empirical = data.frame(a = c(), b = c())

set.seed(11381)
a_star = runif(1000)
M_vector = c()

for (i in 1:1000) {
    y = runif(1000)
    u = runif(1000)
    copula_cond = function(b) {
        copula_pdf(a_star[i],b)
    }
    M = optimize(copula_cond, c(0,1), maximum = T)$obj+0.00001
    M_vector = c(M_vector,M)
    b_star = y[u<=copula_cond(y)/M]
    empirical = rbind(empirical, c(a_star[i],b_star[1]))
}
names(empirical) = c("a","b")
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
plot(empirical$a, empirical$b, xlab = "x", ylab = "y")
```

*(x,y)* 的散佈圖看起來右上及左下資料點較為集中，以下繪製二維的 density plot 較為容易觀察

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(empirical, aes(a,b)) + 
    stat_density2d_filled() + 
    labs(x = "x", y = "y")
```

可以觀察到抽出的樣本在二維空間的分布和 target pdf 有許多特徵相似：右上和左下數據較為集中且以右上最多，圖形對稱於*y=x*

```{r echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow = c(1,2))
hist(empirical$a, prob = T, xlab = "x", main = "marginal pdf")
abline(h = 1, lwd = 2, col = 2)
hist(empirical$b, prob = T, xlab = "y", main = "marginal pdf")
abline(h = 1, lwd = 2, col = 2)
```

*x* 和 *y* 的 marginal empirical pdf 和 *U(0,1)* 相當接近，也符合 *f(x,y)* 是一個 *copula pdf* 的特性。

```{r message=FALSE, warning=FALSE, include=FALSE}
M_inverve = 1/M_vector
mean(M_inverve)
```
Theoretical average efficiency :
\[\frac{1}{10000}\sum_{i=1}^{10000}\frac{1}{M_i}\ \approx\ 62.97\%\]

```{r}
```


```{r}
```

















