---
title: "品質管制 Homework 10"
header-includes:
  - \usepackage{xeCJK}
  - \setCJKmainfont{標楷體}
output: 
  pdf_document:
    latex_engine: xelatex
    number_sections: true
author: "110024516 統研碩一 邱繼賢"
---
**5.13**  
Compute the charting statistics 
$$E_n\ =\ \lambda\ (\frac{X_n-\mu_0}{\sigma_0})^2\ +\ (1-\lambda)E_{n-1}\ ,\ \text{where}\ E_0\ =\ 1\ ,\ \lambda\ =\ 0.1\ ,\ (\mu_0,\sigma_0)=(50,5)$$
A signal of upward process variance shift is given at the n-th time point if 
$$E_n\ >\ U\ =\ 1\ +\ \rho_U\ \sqrt{\frac{2\lambda}{2-\lambda}[1-(1-\lambda)^{2n}]}$$
where $ARL_0=200$ and $\lambda=0.1$ by Table 5.4 $\Rightarrow\ \rho_U=2.595$  
The charting statistics, control limits, and the EWMA chart are shown as below.

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(latex2exp)
x = c(33,55,51,57,57,56,49,45,55,50,
      50,49,50,55,57,50,47,39,51,55,
      52,54,36,37,43,57,52,49,51,47,
      40,35,36,51,39,44,35,54,38,47)
En = c()
En[1] = 0.1*((x[1]-50)/5)^2+0.9*1
for (i in 2:40) {
    En[i] = 0.1*((x[i]-50)/5)^2+0.9*En[i-1]
}

rho_u = 2.595
u = c()
for (i in 1:40) {
    u[i] = 1+rho_u*sqrt((2*0.1/(2-0.1))*(1-(1-0.1)^(2*i)))
}

kable(data.frame(a = 1:40, 
                 b = x, 
                 c = En, 
                 d = u), col.names = c("n", "$X_n$","$E_n$","U"), digits = 4)
```


```{r echo=FALSE}
ii = seq(1,40)
par(mfrow = c(1,1), mar = c(3.5,3.5,2,2))
plot(ii, En, type="o",lty=1,pch=16,xlab="n",ylab=expression(E[n]),
     mgp=c(2,1,0),cex = 0.8, xlim = c(0,40), ylim = c(0.5,4))
lines(ii, u, lty = 2, cex = 0.8)
axis(4, at = u[40], labels = "U", las = 2)
```

We detect the upward variance shifts at the 1st to 6th, 24th to 27th, and time points after 32nd.


\newpage
**5.14**  
Compute the charting statistics 
$$E_n\ =\ \lambda\ (\frac{X_n-\mu_0}{\sigma_0})^2\ +\ (1-\lambda)E_{n-1}\ ,\ \text{where}\ E_0\ =\ 1\ ,\ \lambda\ =\ 0.05\ ,\ (\mu_0,\sigma_0)=(0,1)$$
A signal of upward(downward) process variance shift is given at the n-th time point if 
$$
\begin{aligned}
&E_n\ >\ U\ =\ 1\ +\ \rho_U\ \sqrt{\frac{2\lambda}{2-\lambda}[1-(1-\lambda)^{2n}]}\\
\text{or}\ &E_n\ <\ L\ =\ 1\ -\ \rho_L\ \sqrt{\frac{2\lambda}{2-\lambda}[1-(1-\lambda)^{2n}]}
\end{aligned}
$$
where $ARL_0=200$ and $\lambda=0.05$ by Table 5.4 $\Rightarrow\ (\rho_U,\rho_L)=(2.017,1.510)$  
The charting statistics, control limits, and the EWMA charts are shown as below.

```{r echo=FALSE, message=FALSE, warning=FALSE}
data = read.table("example55.dat.txt")
colnames(data) = c("X1", "X2")

En1 = c()
En2 = c()
En1[1] = 0.05*data$X1[1]^2 + 0.95*1
En2[1] = 0.05*data$X2[1]^2 + 0.95*1

for (i in 2:100) {
    En1[i] = 0.05*data$X1[i]^2 + 0.95*En1[i-1]
    En2[i] = 0.05*data$X2[i]^2 + 0.95*En2[i-1]
}
rho_u = 2.017
rho_l = 1.51
u = c()
l = c()
for (i in 1:100) {
    u[i] = 1+rho_u*sqrt((2*0.05/(2-0.05))*(1-0.95^(2*i)))
    l[i] = 1-rho_l*sqrt((2*0.05/(2-0.05))*(1-0.95^(2*i)))
}
kable(data.frame(a=1:100, 
                 x1 = data$X1, 
                 E1 = En1, 
                 x2 = data$X2, 
                 E2 = En2, 
                 u = u, 
                 l = l), 
      col.names = c("n","$X_{n,1}$", "$E_{n,1}$","$X_{n,2}$","$E_{n,2}$","U", "L"), 
      digits = 4)
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
ii = 1:100
par(mfrow = c(1,2))
plot(ii, En1, type="o",lty=1,pch=16,xlab="n",ylab=expression(E1[n]),
     mgp=c(2,1,0),cex = 0.8, ylim = c(0.2,4))
lines(ii, u, lty = 2, cex = 0.8)
lines(ii, l, lty = 2, cex = 0.8)
axis(4, at = c(u[100],l[100]), labels = c("U","L"), las = 2)

plot(ii, En2, type="o",lty=1,pch=16,xlab="n",ylab=expression(E2[n]),
     mgp=c(2,1,0),cex = 0.8, ylim = c(0.2,1.5))
lines(ii, u, lty = 2, cex = 0.8)
lines(ii, l, lty = 2, cex = 0.8)
axis(4, at = c(u[100],l[100]), labels = c("U","L"), las = 2)
```

For the first dataset, we detect the upward variance shift at the 54th time point, and for the second dataset, we detect the downward shift at the 64th time point.


\newpage
**5.17**  
For detecting the variance upward shifts, compute the charting statistics
$$E_{n,v}\ =\ \lambda\ \frac{S_n^2}{\sigma_0^2}\ +\ (1-\lambda)\ E_{n-1,v}\ \text{, where }E_{0,v}=1\ ,\ \lambda=0.1\ ,\ \sigma_0=0.5$$
A signal of upward process variance shift is given at the n-th time point if
$$E_{n,v}\ >\ U\ =\ 1+\rho_U\ \sqrt{\frac{2\lambda}{(2-\lambda)(m-1)}[1-(1-\lambda)^{2n}]}\ \text{, where }\rho_U\ =\ 2.836\ \text{by Exercise 5.16}$$
For detecting the mean shift, compute the charting statistics
$$E_{n,m}\ =\ \lambda\ \overline{X}_n\ +\ (1-\lambda)\ E_{n-1,m}\ \text{, where }E_0\ =\ \mu_0\ =\ 1$$
The control limits of the EWMA chart for detecting a mean shift is given by
$$
\begin{aligned}
&U\ =\ \mu_0\ +\ \rho\sqrt{\frac{\lambda}{2-\lambda}\ [1-(1-\lambda)^{2n}]}\ \frac{\sigma_0}{\sqrt{m}}\\
&C\ =\ \mu_0\\
&L\ =\ \mu_0\ -\ \rho\sqrt{\frac{\lambda}{2-\lambda}\ [1-(1-\lambda)^{2n}]}\ \frac{\sigma_0}{\sqrt{m}}
\end{aligned}
$$
The charting statistics, control limits, and the EWMA charts are shown as below.

```{r echo=FALSE}
batch = read.table("ex415.dat.txt", header = T)
En_v = c()
En_m = c()
En_v[1] = 0.1*batch$xstd[1]^2/0.25+0.9*1
En_m[1] = 0.1*batch$xbar[1] + 0.9*1
for (i in 2:24) {
    En_v[i] = 0.1*batch$xstd[i]^2/0.25 + 0.9*En_v[i-1]
    En_m[i] = 0.1*batch$xbar[i] + 0.9*En_m[i-1]
}
U_v = c()
u = c()
l = c()
c = 1
for (i in 1:24) {
    U_v[i] = 1+2.836*sqrt((2*0.1/((2-0.1)*4))*(1-0.9^(2*i))^2)
    u[i] = 1+2.731*sqrt((0.1/(2-0.1))*(1-0.9^(2*i)))
    l[i] = 1-2.731*sqrt((0.1/(2-0.1))*(1-0.9^(2*i)))
}
df = data.frame(xbar = batch$xbar, 
                sd = batch$xstd, 
                Env = En_v, 
                Uv = U_v, 
                Enm = En_m, 
                u = u, 
                c = c, 
                l = l)
kable(df, 
      col.names = c("$\\bar{X}_n$", "$S_n$", "$E_{n,v}$","$U_v$","$E_{n,m}$","$U_m$", "$C_m$","$L_m$"), digits = 4)
```


```{r echo=FALSE}
par(mfrow = c(1,2), mar=c(5,4,4,3))
plot(1:24, En_v, type="o",lty=1,pch=16,xlab="n",ylab=expression(E["n,v"]),
     mgp=c(2,1,0),cex = 0.8)
lines(1:24, U_v, lty = 2, cex = 0.8)
axis(4, at = U_v[24], labels = "U", las = 2)

plot(1:24, En_m, type="o",lty=1,pch=16,xlab="n",ylab=expression(E["n,m"]),
     mgp=c(2,1,0),cex = 0.8, ylim = c(0.3,1.7))
lines(1:24, u, lty = 2, cex = 0.8)
lines(1:24, l, lty = 2, cex = 0.8)
lines(1:24, rep(c,24), lty = 2, cex = 0.8)
axis(4, at = c(u[24],l[24], c), labels = c("U","L","C=1"), las = 2)
```

We detect the upward variance shift at the 14th time point, detect the mean shift at the 24th time point.

```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```

