---
title: "Experimental Design and Analysis Homework 5"
author: "110024516 邱繼賢"
header-includes:
  - \usepackage{xeCJK}
  - \setCJKmainfont{標楷體}
  - \linespread{1.5}
output: 
  pdf_document:
    latex_engine: xelatex
---

## Problem 1. (3-26)  
此實驗總共有：  

(1) 四個 two-level treatment factors，分別為 *shield, knob a, knob b, knob c*  
(2) 一個 four-level block factor *day* with block size = 16  

因此我們可以將全部 $2^4$ level combinations 安排進每一天 (block) 的實驗之中，但是本實驗主要的目的是探討哪種 *knobs (a,b,c)* 的設定組合在兩種不同的 *shield* 可以造成最大的 difference，考慮到日照強度會隨著時間有顯著變化，在相同的 *knobs* level combination 下，搭配兩種 *shield* 所執行的兩次實驗應該安排為連續進行，故實驗的設計可以將 30 分鐘為一單位，將一天分為 8 等分，把 $2^3$ 種 *knobs (a,b,c)* 的 level combinations 隨機安排進去，然後再將 *shield* I & II 以各 15 分鐘分配給八個 *knobs* 所形成的 level combinations，在相同的 *knobs* 設定值之下，*shield* I & II 要以 balanced randomize 的方式安排進入實驗的四天 (隨機的兩天先 I 後 II，剩餘兩天先 II 後 I)  

此實驗設計為一個 Split-plot design :  

(1) 3 whole-plot factors : *knob a, knob b, knob c*  
(2) 1 sub-plot factor : *shield*  
(3) 1 block factor : *day*  
(4) whole-plot experimental unit : 30 min  
(5) sub-plot experimental unit : 15 min  


\newpage
level combinations for whole-plot factors :  

| **knob a** | **knob b** | **knob c** | **level combination** |
|:----------:|:----------:|:----------:|:---------------------:|
| +          | +          | +          | 1                     |
| +          | +          | -          | 2                     |
| +          | -          | +          | 3                     |
| +          | -          | -          | 4                     |
| -          | +          | +          | 5                     |
| -          | +          | -          | 6                     |
| -          | -          | +          | 7                     |
| -          | -          | -          | 8                     |

Design plan of four successive days :  

| **Run** | **Day 1**     | **Day 2**     | **Day 3**     | **Day 4**     |
|:-------:|:-------------:|:-------------:|:-------------:|:-------------:|
| 1       | (1,I)  (1,II) | (1,I)  (1,II) | (1,II)  (1,I) | (1,II)  (1,I) |
| 2       | (2,I)  (2,II) | (2,II)  (2,I) | (2,II)  (2,I) | (2,I)  (2,II) |
| 3       | (3,I)  (3,II) | (3,II)  (3,I) | (3,I)  (3,II) | (3,II)  (3,I) |
| 4       | (4,II)  (4,I) | (4,II)  (4,I) | (4,I)  (4,II) | (4,I)  (4,II) |
| 5       | (5,II)  (5,I) | (5,I)  (5,II) | (5,I)  (5,II) | (5,II)  (5,I) |
| 6       | (6,I)  (6,II) | (6,II)  (6,I) | (6,II)  (6,I) | (6,I)  (6,II) |
| 7       | (7,II)  (7,I) | (7,I)  (7,II) | (7,II)  (7,I) | (7,I)  (7,II) |
| 8       | (8,I)  (8,II) | (8,I)  (8,II) | (8,II)  (8,I) | (8,II)  (8,I) |

每天實驗進行的順序應再進行 randomize，此處僅為展示方便  

建構模型以供分析：
$$
y\ =\ \eta\ +\ \tau\ +\ \alpha\ +\ \epsilon^W\ +\ \beta\ +\ (\alpha\beta)\ +\ \epsilon^S
$$
Terms related to the whole-plot :  

(1) $\tau$ : block (*day*) effect  
(2) $\alpha$ : all factorial effects of whole-plot factors (*knob a,b,c*)  
(3) $\epsilon^W$ : whole-plot error  

Terms related to the sub-plot :  

(1) $\beta$ : all factorial effects of sub-plot factor (*shield*)  
(2) $(\alpha\beta)$ : all interaction effects of whole-plot and sub-plot factors (*knob a,b,c* $\times$ *shield*)  
(3) $\epsilon^S$ : sub-plot error  




\newpage
## Problem 3. (4-11)  

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(knitr)
task_data = read.table("TaskEfficiency.txt")
colnames(task_data) = c("setup", "flasher", "inertia", "task", "time", "order")
kable(task_data)
```


### (a)  
Main effect of task
$$
\begin{aligned}
&ME(\text{task})\ =\ \bar{z}(\text{task}=Y)\ -\ \bar{z}(\text{task}=Z)\\
=\ &\frac{1}{4}(11+12+10+11)\ -\ \frac{1}{4}(16+14+15+19)\ =\ -5
\end{aligned}
$$

```{r}
(11+12+10+11)/4 - (16+14+15+19)/4
```

### (b)  
Construct linear model for three treatment factors : *flasher (A) , inertia (B) , task (C)*
$$
\text{time}\ =\ X\beta\ +\ \epsilon
$$
where $X$ is the $2^3$ full factorial design  
Then the factorial effects
$$
\hat{\theta}\ =\ 2\hat{\beta}
$$

```{r}
options(contrasts=c("contr.sum","contr.poly"))
mod3.1 = lm(time ~ flasher*inertia*task, task_data)
factorial_effect = coef(mod3.1)[-1]*2
names(factorial_effect) = c("A","B","C","AB","AC","BC","ABC")
"halfnorm" <- function (x, nlab = 2, labs = as.character(1:length(x)), ylab = "Sorted Data") {
    x <- abs(x)
    labord <- order(x)
    x <- sort(x)
    i <- order(x)
    n <- length(x)
    ui <- qnorm((n + 1:n)/(2 * n + 1))
    plot(ui, x[i], xlab = "Half-normal quantiles", ylab = ylab, ylim=c(0,max(x)),
         type = "n")
    if(nlab < n)
        points(ui[1:(n - nlab)], x[i][1:(n - nlab)])
    text(ui[(n - nlab + 1):n], x[i][(n - nlab + 1):n], labs[labord][(n - nlab + 1):n])
}
halfnorm(factorial_effect, nlab = length(factorial_effect), 
         labs = names(factorial_effect), ylab = "Absolute Effect")
```

By the half-normal plot above, we can see that only the main effect of factor *task (C)* is relatively significant.  

Let's try Lenth's method
$$
PSE\ =\ 1.5\times \text{median}_{\left\{|\hat{\theta}_i|<2.5s_0\right\}}|\hat{\theta}_i|\ =\ 2.5
$$
where
$$
s_0\ =\ 1.5\times\ \text{median}|\hat{\theta}_i|\ =\ 2.5
$$
```{r}
s0 = 1.5*median(abs(factorial_effect))
pse = 1.5*median(abs(factorial_effect[abs(factorial_effect)<2.5*s0]))
```

Compute test statistic
$$
t_{PSE,i}\ =\ \frac{\hat{\theta}_i}{PSE}
$$

```{r}
round(factorial_effect/pse, 4)
```

If $|t_{PSE,i}|$ exceeds the critical value $IER_{0.1}\ =\ 1.71$, we conclude that factorial effect is significant.  
The main effect of factor *task* is significant under Lenth's method. The result is the same with the half-normal plot.  


### (c)  
$$
\begin{aligned}
&\hat{\theta}_C\ =\ \frac{1}{4}(z_1+z_2+z_3+z_4)\ -\ \frac{1}{4}(z_5+z_6+z_7+z_8)\\
\Rightarrow\ &Var\left(\hat{\theta}_C\right)\ =\ \frac{1}{16}\sum_{i=1}^8Var(z_i)\ =\ \frac{1}{16}(4\times1^2\ +\ 4\times4^2)\ =\ 4.25
\end{aligned}
$$





## Problem 4. (4-14)  
### (a)  
No, they can't. Because the effects of width and filler are totally aliasing.  

### (b)  
They can do the $2^2$ factorial design as below design matrix  

| Run | Part Width | Filler |
|:---:|:----------:|:------:|
| 1   | 36         | 40     |
| 2   | 50         | 20     |
| 3   | 36         | 20     |
| 4   | 50         | 40     |

which is an orthogonal array with strength 2  


### (c)  
They have to add more than 2 additional runs if they want to estimate all main and interaction effects. Namely construct a $2^3$ full factorial design whose design matrix is an orthogonal array with strength 3.  

| Run | Part Width | Filler | Temperature |
|:---:|:----------:|:------:|:-----------:|
| 1   | 36         | 40     | 20          |
| 2   | 50         | 20     | 20          |
| 3   | 36         | 20     | 40          |
| 4   | 50         | 40     | 40          |
| 5   | 36         | 40     | 20          |
| 6   | 50         | 20     | 20          |
| 7   | 36         | 20     | 40          |
| 8   | 50         | 40     | 40          |


However, if we assume that the interaction among the three factors can be neglected and estimating only all main effects is their purpose, they can just add 2 additional runs in the design plan.  

| Run | Part Width | Filler | Temperature |
|:---:|:----------:|:------:|:-----------:|
| 1   | 36         | 40     | 20          |
| 2   | 50         | 20     | 20          |
| 3   | 36         | 20     | 40          |
| 4   | 50         | 40     | 40          |

which is still an orthogonal array with strength 2, and it is enough to estimate the main effects.  


## Problem 5. (4-16)  
### (a)  
Construct location model and dispersion model
$$
\bar{y}\ =\ X\beta\ +\ \epsilon\ \ ,\ \ \ln s^2\ =\ X\gamma\ +\ \delta
$$
where $X$ is the $2^3$ full factorial design.  

Then the factorial effects for location and dispersion effects are
$$
\hat{\theta}\ =\ -2\hat{\beta}\ \ ,\ \ \hat{\psi}\ =\ -2\hat{\gamma}
$$
the negative sign is because the signs of sum coding in model matrix and $(-,+)$ levels for the factors are just the opposite.  

```{r}
drive_data = read.table("DriveShaft.txt")
colnames(drive_data) = c("run", "A", "B", "C", "y1", "y2", "y3", "y4", "y5")
y_bar = apply(drive_data[,5:9],1,mean)
s_square = apply(drive_data[,5:9],1,var)
drive_data = drive_data %>% mutate(y_bar, s_square)
options(contrasts=c("contr.sum","contr.poly"))
mod_loc = lm(y_bar ~ A*B*C, drive_data)
mod_dis = lm(log(s_square) ~ A*B*C, drive_data)
loc_effect = -coef(mod_loc)[-1]*2 ; names(loc_effect) = c("A","B","C","AB","AC","BC","ABC")
loc_effect
dis_effect = -coef(mod_dis)[-1]*2 ; names(dis_effect) = c("A","B","C","AB","AC","BC","ABC")
round(dis_effect,4)
```


Let's check the significance of location and dispersion effects :  

(1) Half-normal plot method  


```{r}
library(latex2exp)
par(mfrow = c(1,2))
halfnorm(loc_effect, nlab = 3, labs = c("A","B","C","AB","AC","BC","ABC"), 
         ylab = "Absolute Effect")
title(TeX("$\\bar{y}$"))
halfnorm(dis_effect, nlab = 3, labs = c("A","B","C","AB","AC","BC","ABC"), 
         ylab = "Absolute Effect")
title(TeX("$ln\ s^2$"))
```

For $\bar{y}$, main effects of B and C are significant. For $\ln s^2$, only the main effect of A looks slightly significant.  

(2) Lenth's method  
Following the same step in **(4-11)** to compute the test statistics $\frac{\hat{\theta}}{PSE_{\hat{\theta}}}$ and $\frac{\hat{\psi}}{PSE_{\hat{\psi}}}$

```{r}
s0_loc = 1.5*median(abs(loc_effect))
pse_loc = 1.5*median(abs(loc_effect[abs(loc_effect)<2.5*s0_loc]))
round(loc_effect/pse_loc,4)
```


```{r}
s0_dis = 1.5*median(abs(dis_effect))
pse_dis = 1.5*median(abs(dis_effect[abs(dis_effect)<2.5*s0_dis]))
round(dis_effect/pse_dis,4)
```

If $|t_{PSE,i}|$ exceeds the critical value $IER_{0.1}\ =\ 1.71$, we conclude that factorial effect is significant.  
For $\bar{y}$, the main effects of B and C are significant. For $\ln s^2$, no effects look significant, only the test statistic of the main effect A slightly closed to the critical value.  


### (b)  
Fit the location model and dispersion model
$$
\begin{aligned}
&\hat{\bar{y}}\ =\ \hat{\beta}_0\ +\ \frac{1}{2}\hat{\theta}_Bx_B\ +\ \frac{1}{2}\hat{\theta}_Cx_C\ =\ 65.42\ -\ 6.8x_B\ +\ 7.3x_C\\
&\ln \hat{s}^2\ =\ \hat{\gamma}_0\ +\ \frac{1}{2}\hat{\psi}_Ax_A\ =\ 3.8714225\ -\ 0.9241892x_A
\end{aligned}
$$
where
$$
\left\{
\begin{aligned}
&x_A\ =\ -1\ \ ,\ \ \text{if}\ \ A\ =\ \#5023\\
&x_A\ =\ \ \ \ \ \ 1\ \ ,\ \ \text{if}\ \ A\ =\ \#5074
\end{aligned}
\right.\ \ ,\ \ 
\left\{
\begin{aligned}
&x_B\ =\ -1\ \ ,\ \ \text{if}\ \ B\ =\ 800\\
&x_B\ =\ \ \ \ \ \ 1\ \ ,\ \ \text{if}\ \ B\ =\ 1000
\end{aligned}
\right.\ \ ,\ \ 
\left\{
\begin{aligned}
&x_C\ =\ -1\ \ ,\ \ \text{if}\ \ C\ =\ 50\\
&x_C\ =\ \ \ \ \ \ 1\ \ ,\ \ \text{if}\ \ C\ =\ 80
\end{aligned}
\right.
$$

It is appropriate to use the two-step procedure because there exist two adjustment factors B and C.  

(1) Choose $A\ =\ \#5074$ to minimize $Var(y_x)$, then the predicted variance
$$
\hat{\sigma}^2\ =\ \exp\left[3.8714225\ -\ 0.9241892(1)\right]\ =\ 19.05317
$$

(2) Choose $x_B$ and $x_C$ to satisfy $75\ =\ 65.42\ -\ 6.8x_B\ +\ 7.3x_C$. For example
$$
\left(x_B\ ,\ x_C\right)\ =\ (-1\ ,\ 0.3808219)\ \Leftrightarrow\ \left(B\ ,\ C\right)\ =\ (800\ ,\ 70.71233)
$$








## Problem 6. (4-20)  
The first blocking scheme $B_{I}$
$$
\begin{aligned}
&B_1=126\ ,\ B_2=136\ ,\ B_3=346\ ,\ B_4=456\ ,\ B_{12}=23\ ,\ B_{13}=1234\ ,\ B_{14}=1245\ ,\\
&B_{23}=14\ ,\ B_{24}=1345\ ,\ B_{34}=35\ ,\ B_{123}=246\ ,\ B_{124}=23456\ ,\ B_{134}=12356\ ,\ B_{234}=156\ ,\ B_{1234}=25\\
\Rightarrow\ &g\left(B_I\right)\ =\ (0,4,6,3,2,0)
\end{aligned}
$$

The second blocking scheme $B_{II}$
$$
\begin{aligned}
&B_1=136\ ,\ B_2=1234\ ,\ B_3=3456\ ,\ B_4=123456\ ,\ B_{12}=246\ ,\ B_{13}=145\ ,\ B_{14}=245\ ,\\
&B_{23}=1256\ ,\ B_{24}=56\ ,\ B_{34}=12\ ,\ B_{123}=235\ ,\ B_{124}=135\ ,\ B_{134}=236\ ,\ B_{234}=34\ ,\ B_{1234}=146\\
\Rightarrow\ &g\left(B_{II}\right)\ =\ (0,3,8,3,0,1)
\end{aligned}
$$

Notice that $g_2(B_I)\ >\ g_2(B_{II})$, so scheme $B_{II}$ is said to have less aberration than scheme $B_I$. It is clear that scheme $B_{II}$ sacrifies less number of two-factor interaction than scheme $B_I$. By the effect hierarchy principle, the second scheme $B_{II}$ is better.  


